### 2016-10-20 methodology.tex

\section{Pruning Neurons to Shrink Neural Networks}\label{sec2}
As discussed in Section \ref{sec1} our aim here is to leverage the non-uniform distribution of learning representation in pre-trained neural networks to eliminate non-contributing neurons in their entirety, without focusing on individual weight parameters. Taking this approach enables us to remove all the weights (incoming and outgoing) associated with a non-contributing neuron in one go. We would like to note here that in an ideal scenario, based on the neuron interdependency theory put forward by \cite{mozer1989skeletonization}, one would evaluate all possible combinations of neurons to remove (one at a time, two at a time, three at a time and so forth) to find the optimal subset of neurons to keep. This however is not feasible practically, which is why we will only focus on removing one neuron at a time and explore "greedy" algorithms to do this in an efficient manner.

The general approach taken to prune an optimally trained neural network here is to create a ranked list of all the neurons in the network based off of one of the 3 proposed ranking criteria: Brute Force approximation (which we use as our ground truth), linear approximation and quadratic approximation of the neuron's impact on the overall performance of the network. We then test the effects of removing neurons on the accuracy and error of the network. All the algorithms and methods presented here are easily parallelizable.

One last thing to note here before moving forward is that the methods discussed in this section involve some non-trivial mathematics, discussing all of which is beyond the scope of this paper as we wish to keep the math to a minimum in the main body of the paper. However, a complete step-by-step derivation and proof of all the results presented is provided in the Supplementary Material as an Appendix.

\subsection{Brute Force Removal Approach}
This is perhaps the most naive yet the most accurate method for pruning the network. It is also the slowest and hence possibly unusable on large-scale neural networks with thousands of neurons. This method explicitly evaluates each neuron in the network. The idea is to manually check the effect of every single neuron on the output. This is done by running a forward propagation on the validation set $K$ times (where $K$ is the total number of neurons in the network), turning off exactly one neuron each time (keeping all other neurons active) and noting down the change in error. Turning a neuron off can be achieved by simply setting its output to 0. This results in all the outgoing weights from that neuron being turned off. This change in error is then used to generate the ranked list.

\input{taylor_series.tex}

\input{less_math_error_approximation.tex}
\input{pruning_algorithm.tex}