### 2017-01-15 iclr2017_conference.bib.tex
arxiv submission

@article{nielsen2015neural,
  title={Neural networks and deep learning},
  author={Nielsen, Michael A},
  journal={URL: http://neuralnetworksanddeeplearning.com/chap4.html.(visited: 01.11. 2014)},
  year={2015}
}
@article{mozer1989using,
  title={Using relevance to reduce network size automatically},
  author={Mozer, Michael C and Smolensky, Paul},
  journal={Connection Science},
  volume={1},
  number={1},
  pages={3--16},
  year={1989},
  publisher={Taylor \& Francis}
}
@article{balzer1991weight,
  title={Weight quantization in Boltzmann machines},
  author={Balzer, Wolfgang and Takahashi, Masanobu and Ohta, Jun and Kyuma, Kazuo},
  journal={Neural Networks},
  volume={4},
  number={3},
  pages={405--409},
  year={1991},
  publisher={Elsevier}
}
@article{dundar1994effects,
  title={The effects of quantization on multilayer neural networks.},
  author={Dundar, Gunhan and Rose, Kenneth},
  journal={IEEE transactions on neural networks/a publication of the IEEE Neural Networks Council},
  volume={6},
  number={6},
  pages={1446--1451},
  year={1994}
}
@article{hoehfeld1992learning,
  title={Learning with limited numerical precision using the cascade-correlation algorithm},
  author={Hoehfeld, Markus and Fahlman, Scott E},
  journal={IEEE Transactions on Neural Networks},
  volume={3},
  number={4},
  pages={602--611},
  year={1992},
  publisher={IEEE}
}
@article{fahlman1989cascade,
  title={The cascade-correlation learning architecture},
  author={Fahlman, Scott E and Lebiere, Christian},
  year={1989}
}
@incollection{chauvin1990generalization,
  title={Generalization performance of overtrained back-propagation networks},
  author={Chauvin, Yves},
  booktitle={Neural Networks},
  pages={45--55},
  year={1990},
  publisher={Springer}
}
@article{reed1993pruning,
  title={Pruning algorithms-a survey},
  author={Reed, Russell},
  journal={Neural Networks, IEEE Transactions on},
  volume={4},
  number={5},
  pages={740--747},
  year={1993},
  publisher={IEEE}
}
@book{hassibi1993second,
  title={Second order derivatives for network pruning: Optimal brain surgeon},
  author={Hassibi, Babak and Stork, David G},
  year={1993},
  publisher={Morgan Kaufmann}
}
@inproceedings{mozer1989skeletonization,
  title={Skeletonization: A technique for trimming the fat from a network via relevance assessment},
  author={Mozer, Michael C and Smolensky, Paul},
  booktitle={Advances in neural information processing systems},
  pages={107--115},
  year={1989}
}
@inproceedings{lecun1989optimal,
  title={Optimal brain damage.},
  author={LeCun, Yann and Denker, John S and Solla, Sara A and Howard, Richard E and Jackel, Lawrence D},
  booktitle={NIPs},
  volume={89},
  year={1989}
}
@incollection{Bengio+chapter2007,
author = {Bengio, Yoshua and LeCun, Yann},
booktitle = {Large Scale Kernel Machines},
publisher = {MIT Press},
title = {Scaling Learning Algorithms Towards {AI}},
year = {2007}
}
@inproceedings{becker1988improving,
  title={Improving the convergence of back-propagation learning with second order methods},
  author={Becker, Sue and Le Cun, Yann},
  booktitle={Proceedings of the 1988 connectionist models summer school},
  pages={29--37},
  year={1988},
}
@inproceedings{segee1991fault,
  title={Fault tolerance of pruned multilayer networks},
  author={Segee, Bruce E and Carter, Michael J},
  booktitle={Neural Networks, 1991., IJCNN-91-Seattle International Joint Conference on},
  volume={2},
  pages={447--452},
  year={1991},
  organization={IEEE}
}
@article{srivastava2014dropout,
  title={Dropout: A simple way to prevent neural networks from overfitting},
  author={Srivastava, Nitish and Hinton, Geoffrey and Krizhevsky, Alex and Sutskever, Ilya and Salakhutdinov, Ruslan},
  journal={The Journal of Machine Learning Research},
  volume={15},
  number={1},
  pages={1929--1958},
  year={2014},
}
@article{goodfellow2013maxout,
  title={Maxout networks},
  author={Goodfellow, Ian J and Warde-Farley, David and Mirza, Mehdi and Courville, Aaron and Bengio, Yoshua},
  journal={arXiv preprint arXiv:1302.4389},
  year={2013}
}
@article{castellano1997iterative,
  title={An iterative pruning algorithm for feedforward neural networks},
  author={Castellano, Giovanna and Fanelli, Anna Maria and Pelillo, Marcello},
  journal={Neural Networks, IEEE Transactions on},
  volume={8},
  number={3},
  pages={519--531},
  year={1997},
  publisher={IEEE}
}
@inproceedings{sietsma1988neural,
  title={Neural net pruning-why and how},
  author={Sietsma, Jocelyn and Dow, Robert JF},
  booktitle={Neural Networks, 1988., IEEE International Conference on},
  pages={325--333},
  year={1988},
  organization={IEEE}
}
@article{baum1989size,
  title={What size net gives valid generalization?},
  author={Baum, Eric B and Haussler, David},
  journal={Neural computation},
  volume={1},
  number={1},
  pages={151--160},
  year={1989},
  publisher={MIT Press}
}
@inproceedings{Anders2016quant,
  author    = {Anders {\O}land and
               Bhiksha Raj},
  title     = {Reducing communication overhead in distributed learning by an order
               of magnitude (almost)},
  booktitle = {{IEEE} International Conference on Acoustics, Speech and Signal
               Processing},
  pages     = {2219--2223},
  year      = {2015}
}
@article{deepcompression2016,
  title={DEEP COMPRESSION: COMPRESSING DEEP NEURAL NETWORKS WITH PRUNING, TRAINED QUANTIZATION AND HUFFMAN CODING},
  author={Han, Song and Mao, Huizi and Dally, William J},
  journal={arXiv preprint arXiv:1510.00149v5},
  year={2016}
}
@article{lecun-mnisthandwrittendigit-2010,
  author = {LeCun, Yann and Cortes, Corinna},
  howpublished = {http://yann.lecun.com/exdb/mnist/},
  title = {{MNIST} handwritten digit database},
  year = {2010}
}
@article{darkknowledge2015,
  title={Distilling the Knowledge in a Neural Network},
  author={Hinton, Geoffrey and Vinyals, Oriol and Dean, Jeff},
  journal={arXiv preprint arXiv:1503.02531},
  year={2015}
}
@inproceedings{prabhavalkar2016compression,
  title={On the compression of recurrent neural networks with an application to LVCSR acoustic modeling for embedded speech recognition},
  author={Prabhavalkar, Rohit and Alsharif, Ouais and Bruguier, Antoine and McGraw, Lan},
  booktitle={2016 IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)},
  pages={5970--5974},
  year={2016},
  organization={IEEE}
}